---
title: "enhancer"
output: html_document
date: "2025-02-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## read
```{r cars}
enhancer_hepg2 <- read.csv("C:/Users/ruini/Desktop/project2/r/motif_info_thresh_500_enhancers_hepg2.csv", header = TRUE, sep = ",")
prediction_enhancer <-read.csv("C:/Users/ruini/Desktop/project2/r/predictions_enhancers_hepg2.csv",header = TRUE)
classify <- read.csv("C:/Users/ruini/Desktop/thesis/tf_cooperativity_index_hepg2_pe.csv",header = TRUE)
gradxinp_deepcompre <- read.csv("C:/Users/ruini/Desktop/thesis/gradxinp_enhancers_hepg2.csv", header = TRUE, sep = ",")
gradxinp_enformer <- read.csv("C:/Users/ruini/Desktop/thesis/Enformer_gradxinp.csv", header = TRUE, sep = ",")
motif_ism <- read.csv("C:/Users/ruini/Desktop/thesis/ism_motif200.csv",header = TRUE, sep = ",")
```

## set save path
```{r}
if (requireNamespace("rstudioapi", quietly = TRUE)) {
  current_path <- rstudioapi::getSourceEditorContext()$path
  if (!is.null(current_path)) {
    setwd(dirname(current_path))
    message("Working directory set to script folder: ", dirname(current_path))
  }
}
```


## data preparation
```{r}
# use abs(max) in the gradxinp for motif
enhancer_hepg2$gradxinp_0_max <- sapply(
  strsplit(enhancer_hepg2$gradxinp_0, ":"),
  function(x) max(abs(as.numeric(x)))
)

enhancer_hepg2$gradxinp_0_max_orginal <- sapply(
  strsplit(enhancer_hepg2$gradxinp_0, ":"),
  function(x) max(as.numeric(x))
)

# calculate gradxinp_min
enhancer_hepg2$gradxinp_0_min <- sapply(
  strsplit(enhancer_hepg2$gradxinp_0, ":"),
  function(x) min(abs(as.numeric(x)))
)

enhancer_hepg2$gradxinp_0_min_orginal <- sapply(
  strsplit(enhancer_hepg2$gradxinp_0, ":"),
  function(x) min(as.numeric(x))
)

# use abs(max) in the isa in base level for motif
enhancer_hepg2$isa_0_max <- sapply(
  strsplit(enhancer_hepg2$isa_0, ":"),
  function(x) max(as.numeric(x))
)

# use max in the ism for motif
enhancer_hepg2$ism_0_max <- sapply(
  strsplit(enhancer_hepg2$ism_0, ":"),
  function(x) max(as.numeric(x))
)

# use max in the ism for motif
enhancer_hepg2$ism_0_max_abs <- sapply(
  strsplit(enhancer_hepg2$ism_0, ":"),
  function(x) max(abs(as.numeric(x)))
)

# use abs(isa)
enhancer_hepg2$isa_abs <- abs(enhancer_hepg2$isa_track0)

```

# motif length
```{r}
# calculate motif length
motif_shortest <- merged_enhancer %>%
  mutate(motif_length = end - start) %>%
  arrange(motif_length)  

# select 200 motifs
motif_shortest_diverse <- motif_shortest %>%
  group_by(chromosome) %>%
  slice_head(n = 20) %>%  
  ungroup() %>%
  slice_head(n = 200)     

# save
write.csv(motif_shortest_diverse, "motif_ism_shortest200.csv", row.names = FALSE)


# 读取之前选过的 motif 数据
previous_motifs <- read.csv("C:/Users/ruini/Desktop/thesis/enhancer/motif_ism_shortest200.csv")
motif_remaining <- anti_join(motif_shortest, previous_motifs, by = c("chromosome", "start", "end"))
motif_shortest_2000 <- motif_remaining %>%
  arrange(motif_length) %>%
  slice_head(n = 2000)
write.csv(motif_shortest_2000, "motif_ism_shortest2000_unique.csv", row.names = FALSE)

```

## add prediction
```{r}
merged_enhancer <- merge(enhancer_hepg2, prediction_enhancer, 
                     by.x = "region", 
                     by.y = "X", 
                     all = FALSE)  
colnames(merged_enhancer)[colnames(merged_enhancer) == "X0"] <- "prediction"
# filter prediction >= 0 
merged_enhancer <- merged_enhancer %>%
  filter(prediction >= 0)
```

## base level scatter
```{r}
parse_column <- function(col) {
  strsplit(as.character(col), ":") %>%
    lapply(function(x) as.numeric(x)) %>%
    unlist()
}

gradx <- parse_column(merged_enhancer$gradxinp_0)
ism <- parse_column(merged_enhancer$ism_0)
isa <- parse_column(merged_enhancer$isa_0)

# calculate correlation
cor_grad_isa <- cor(gradx, isa)
cor_grad_ism <- cor(gradx, ism)
cor_isa_ism <- cor(isa, ism)

# theme
theme_scatter <- theme_minimal() +
  theme(
    panel.grid.major = element_line(color = "grey80", size = 0.3),
    panel.grid.minor = element_line(color = "grey90", size = 0.2),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    plot.background = element_rect(fill = "white", color = NA)
  )


# Gradxinp vs ISA
p1 <- ggplot(data.frame(gradx, isa), aes(x = gradx, y = isa)) +
  geom_point(alpha = 0.4, size = 0.3) +
  annotate("text", x = Inf, y = Inf, label = sprintf("r = %.2f", cor_grad_isa),
           hjust = 1.1, vjust = 1.5, size = 4) +
  labs(x = "Gradxinp", y = "ISA") +
  theme_scatter

ggsave("scatter_gradx_vs_isa.pmerged_enhancer.png", p1, width = 6, height = 4)

# Gradxinp vs ISM
p2 <- ggplot(data.frame(gradx, ism), aes(x = gradx, y = ism)) +
  geom_point(alpha = 0.4, size = 0.3) +
  annotate("text", x = Inf, y = Inf, label = sprintf("r = %.2f", cor_grad_ism),
           hjust = 1.1, vjust = 1.5, size = 4) +
  labs(x = "Gradxinp", y = "ISM") +
  theme_scatter

ggsave("scatter_gradx_vs_ism.pmerged_enhancer.png", p2, width = 6, height = 4)

# ISA vs ISM
p3 <- ggplot(data.frame(isa, ism), aes(x = isa, y = ism)) +
  geom_point(alpha = 0.4, size = 0.3) +
  annotate("text", x = Inf, y = Inf, label = sprintf("r = %.2f", cor_isa_ism),
           hjust = 1.1, vjust = 1.5, size = 4) +
  labs(x = "ISA", y = "ISM") +
  theme_scatter

ggsave("scatter_isa_vs_ism.pmerged_enhancer.png", p3, width = 6, height = 4)


```

# add 200 examples of motif ISM
```{r}
motif_ism_merged <- motif_ism %>%
  inner_join(merged_enhancer %>% 
               select(region, start, end, protein, isa_track0, prediction),
             by = c("region", "start", "end", "protein"))

print(head(motif_ism_merged))
```

## function of scatter plot
```{r}
library(ggplot2)
library(ggpubr)

scatter_plot_with_cor <- function(data, x_var, y_var) {
  theme_with_border <- theme_minimal() +
    theme(
      panel.border = element_rect(color = "black", fill = NA, size = 0.8),
      plot.background = element_rect(fill = "white", color = NA),
      axis.title = element_text(size = 16),
      axis.text = element_text(size = 14),
      plot.title = element_text(size = 18, face = "bold"),
      legend.text = element_text(size = 14),
      legend.title = element_text(size = 15)
    )
  
  p <- ggplot(data, aes(x = .data[[x_var]], y = .data[[y_var]])) +
    geom_point(alpha = 0.5) +
    ggpubr::stat_cor(
      method = "pearson",
      label.x.npc = 0.2,
      label.y.npc = 0.9,
      size = 5
    )+ 
    labs(x = x_var, y = y_var) +
    theme_with_border
  
  return(p)
}

```

## motif ism vs motif isa
```{r}
plot <- scatter_plot_with_cor(
  data = motif_ism_merged, 
  x_var = "ism_score",   
  y_var = "isa_track0"   
)
plot
ggsave(
    filename = "motif_ism_vs_isa.png",
    plot = plot,
    width = 8,
    height = 6,
    dpi = 300,
    bg = "white"
  )
```
## max abs gradxinp vs isa on motif level
```{r}
plot <- scatter_plot_with_cor(
  data = merged_enhancer, 
  x_var = "gradxinp_0_max",   
  y_var = "isa_track0"   
)
plot
ggsave(
    filename = "gradxinp_abs_max_vs_isa.png",
    plot = plot,
    width = 8,
    height = 6,
    dpi = 300,
    bg = "white"
  )
```
## gradxinp max vs min
```{r}
plot <- scatter_plot_with_cor(
  data = merged_enhancer, 
  x_var = "gradxinp_0_max_orginal",   
  y_var = "gradxinp_0_min_orginal"   
)
plot
ggsave(
    filename = "gradxinp_max_vs_min.png",
    plot = plot,
    width = 8,
    height = 6,
    dpi = 300,
    bg = "white"
  )
```
# isa vs ism
```{r}
plot <- scatter_plot_with_cor(
  data = merged_enhancer, 
  x_var = "ism_0_max",   
  y_var = "isa_track0"   
)
plot
ggsave(
    filename = "ism_max_vs_isa.png",
    plot = plot,
    width = 8,
    height = 6,
    dpi = 300,
    bg = "white"
  )
```
# isa vs ism_abs
```{r}
plot <- scatter_plot_with_cor(
  data = merged_enhancer, 
  x_var = "ism_0_max_abs",   
  y_var = "isa_track0"   
)
plot
```


# gradxinp vs ism
```{r}
plot <- scatter_plot_with_cor(
  data = merged_enhancer, 
  x_var = "ism_0_max",   
  y_var = "gradxinp_0_max"   
)
plot
ggsave(
    filename = "ism_max_vs_gradxinp_abs_max.png",
    plot = plot,
    width = 8,
    height = 6,
    dpi = 300,
    bg = "white"
  )
```



### predition analysis
# prediction vs isa
```{r}
plot <- scatter_plot_with_cor(
  data = merged_enhancer, 
  x_var = "prediction",   
  y_var = "isa_track0"   
)
plot
ggsave(
    filename = "isa_vs_prediction.png",
    plot = plot,
    width = 8,
    height = 6,
    dpi = 300,
    bg = "white"
  )
```
# prediction vs gradxinp
```{r}
plot <- scatter_plot_with_cor(
  data = merged_enhancer, 
  x_var = "prediction",   
  y_var = "gradxinp_0_max"   
)
plot
ggsave(
    filename = "gradxinp_vs_prediction_all.png",
    plot = plot,
    width = 8,
    height = 6,
    dpi = 300,
    bg = "white"
  )
```
# prediction vs ism
```{r}
plot <- scatter_plot_with_cor(
  data = merged_enhancer, 
  x_var = "prediction",   
  y_var = "ism_0_max"   
)
plot
ggsave(
    filename = "ism_vs_prediction_all.png",
    plot = plot,
    width = 8,
    height = 6,
    dpi = 300,
    bg = "white"
  )
```
# prediction vs motif ism
```{r}
plot <- scatter_plot_with_cor(
  data = motif_ism_merged, 
  x_var = "prediction",   
  y_var = "ism_score"   
)
plot
ggsave(
    filename = "motif_ism_vs_prediction_all.png",
    plot = plot,
    width = 8,
    height = 6,
    dpi = 300,
    bg = "white"
  )
```

## find high ISM and low prediction
```{r}
ism_high_pred_low <- merged_enhancer %>%
  filter(
    ism_0_max > 2,
    prediction < 0.2
  )
ism_high_pred_low
```

## correlation distribution
```{r}
library(ggrepel)  
library(dplyr)

safe_correlation_plot <- function(
    data, 
    protein_col = "protein",
    x_var = "prediction",
    y_var = "gradxinp_0_max",
    bins = 30,
    fill_colors = c("Extreme" = "darkred", "Normal" = "steelblue"),
    label_color = "navy",
    top_percent = 0.03
) {
  
  if (!all(c(protein_col, x_var, y_var) %in% names(data))) {
    stop("Required columns not found in dataset")
  }
  
  # count the protein
  protein_counts <- data %>%
    count(!!sym(protein_col), name = "protein_count")
  
  # calculate correlation
  correlations <- data %>%
    group_by(!!sym(protein_col)) %>%
    summarise(
      correlation = suppressWarnings(
        cor(!!sym(x_var), !!sym(y_var), 
            method = "pearson", use = "complete.obs")
      ),
      .groups = "drop"
    ) %>%
    filter(!is.na(correlation)) %>%
    left_join(protein_counts, by = protein_col) %>%
    filter(!is.na(protein_count), protein_count > 0)
  
  if (nrow(correlations) == 0) {
    stop("No valid correlations calculated")
  }
  
  # calculate threshold
  threshold_low <- quantile(correlations$correlation, top_percent/2)
  threshold_high <- quantile(correlations$correlation, 1 - top_percent/2)
  
  # label
  extreme_proteins <- correlations %>%
    filter(correlation <= threshold_low | correlation >= threshold_high) %>%
    mutate(
      label = paste0(!!sym(protein_col), "(n=", protein_count, ")"),
      y_pos = pmax(pmin(10, protein_count), 1) 
    ) %>%
    filter(!is.na(correlation), is.finite(correlation))
  
  # plot
  p <- ggplot(correlations, aes(x = correlation)) +
    geom_histogram(
      aes(fill = ifelse(
        correlation <= threshold_low | correlation >= threshold_high,
        "Extreme", "Normal")),
      bins = bins,
      color = "white",
      alpha = 0.7
    ) +
    scale_fill_manual(values = fill_colors) +
    labs(
      title = paste("Correlation between", x_var, "and", y_var),
      x = "Pearson Correlation Coefficient",
      y = "Number of Proteins",
      fill = "Category"
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(hjust = 0.5, size = 18),
      axis.title = element_text(size = 16),
      axis.text = element_text(size = 14),
      legend.title = element_text(size = 15),
      legend.text = element_text(size = 14),
      legend.position = "top",
      panel.border = element_rect(color = "black", fill = NA, size = 0.8)
    )

  
  if (nrow(extreme_proteins) > 0) {
    p <- p +
      geom_text_repel(
        data = extreme_proteins,
        aes(x = correlation, y = y_pos, label = label),
        direction = "y",
        angle = 90,
        hjust = 0,
        vjust = 0.5,
        color = label_color,
        size = 3,
        segment.color = "gray",
        box.padding = 0.5,
        force = 0.5,
        max.time = 2,
        nudge_y = 2,
        min.segment.length = 0,
        max.overlaps = Inf
      ) +
      scale_y_continuous(expand = expansion(mult = c(0, 0.95)))
  }
  
  return(p)
}
```

# correlation between prediction and max gradxinp vs count
```{r}
plot <- safe_correlation_plot(merged_enhancer)
plot
ggsave(
    filename = "correlation_gradxinp_max.png",
    plot = plot,
    width = 8,
    height = 6,
    dpi = 300,
    bg = "white"
  )
```
# correlation between prediction and ism vs count
```{r}
plot <- safe_correlation_plot(
  data = merged_enhancer,
  x_var = "prediction",
  y_var = "ism_0_max",
  fill_colors = c("Extreme" = "darkred", "Normal" = "steelblue"),
  label_color = "navy",
  top_percent = 0.03
)
plot
ggsave(
    filename = "correlation_ism_max.png",
    plot = plot,
    width = 8,
    height = 6,
    dpi = 300,
    bg = "white"
  )
```
# correlation between prediction and isa vs count
```{r}
plot <- safe_correlation_plot(
  data = merged_enhancer,
  x_var = "prediction",
  y_var = "isa_track0",
  fill_colors = c("Extreme" = "darkred", "Normal" = "steelblue"),
  label_color = "navy",
  top_percent = 0.03
)
plot
ggsave(
    filename = "correlation_isa.png",
    plot = plot,
    width = 8,
    height = 6,
    dpi = 300,
    bg = "white"
  )
```
# median correlations
```{r}
cor_pre_isa = cor(merged_enhancer$prediction, merged_enhancer$isa_track0, method = "pearson")
cor_pre_isa
cor_pre_grad = cor(merged_enhancer$prediction, merged_enhancer$gradxinp_0_max, method = "pearson")
cor_pre_grad
cor_pre_ism = cor(merged_enhancer$prediction, merged_enhancer$ism_0_max, method = "pearson")
cor_pre_ism
```
# function for plotting
```{r}
scatter_plot_with_cor2 <- function(data, x_var, y_var, protein_name) {
  filtered_data <- data %>% filter(protein == protein_name)

  theme_with_border <- theme_minimal() +
    theme(
      panel.border = element_rect(color = "black", fill = NA, size = 0.8),
      plot.background = element_rect(fill = "white", color = NA),
      axis.title = element_text(size = 16),
      axis.text = element_text(size = 14),
      plot.title = element_text(hjust = 0.5, size = 18),
      legend.text = element_text(size = 14),
      legend.title = element_text(size = 15)
    )

  p <- ggplot(filtered_data, aes(x = .data[[x_var]], y = .data[[y_var]])) +
    geom_point(alpha = 0.5) +
    stat_cor(
      method = "pearson",
      label.x.npc = 0.1,
      label.y.npc = 0.3,
      size = 6
    ) +
    labs(
      title = protein_name,
      x = x_var,
      y = y_var
    ) +
    theme_with_border

  return(p)
}

# remove outliers
scatter_plot_with_cor3 <- function(data, x_var, y_var, protein_name) {
  filtered_data <- data %>% filter(protein == protein_name)
  filtered_data <- filtered_data %>% filter(prediction < 1.5)

  theme_with_border <- theme_minimal() +
    theme(
      panel.border = element_rect(color = "black", fill = NA, size = 0.8),
      plot.background = element_rect(fill = "white", color = NA),
      axis.title = element_text(size = 16),
      axis.text = element_text(size = 14),
      plot.title = element_text(hjust = 0.5, size = 18),
      legend.text = element_text(size = 14),
      legend.title = element_text(size = 15)
    )

  p <- ggplot(filtered_data, aes(x = .data[[x_var]], y = .data[[y_var]])) +
    geom_point(alpha = 0.5) +
    stat_cor(
      method = "pearson",
      label.x.npc = 0.1,
      label.y.npc = 0.1,
      size = 6
    ) +
    labs(
      title = paste0(protein_name, " (removing outliers)"),
      x = x_var,
      y = y_var
    ) +
    theme_with_border

  return(p)
}
```


# check extreme proteins for ISM
```{r}
# TEAD1
plot <- scatter_plot_with_cor2(
  data = merged_enhancer, 
  x_var = "prediction",   
  y_var = "ism_0_max",
  protein_name = "TEAD1"
)
plot
ggsave(
    filename = "ISM_prediction_TEAD1_1.png",
    plot = plot,
    width = 8,
    height = 6,
    dpi = 300,
    bg = "white"
  )
plot <- scatter_plot_with_cor3(
  data = merged_enhancer, 
  x_var = "prediction",   
  y_var = "ism_0_max",
  protein_name = "TEAD1"
)
plot
ggsave(
    filename = "ISM_prediction_TEAD1_2.png",
    plot = plot,
    width = 8,
    height = 6,
    dpi = 300,
    bg = "white"
  )
# TEAD4
plot <- scatter_plot_with_cor2(
  data = merged_enhancer, 
  x_var = "prediction",   
  y_var = "ism_0_max",
  protein_name = "TEAD4"
)
plot
ggsave(
    filename = "ISM_prediction_TEAD4_1.png",
    plot = plot,
    width = 8,
    height = 6,
    dpi = 300,
    bg = "white"
  )
plot <- scatter_plot_with_cor3(
  data = merged_enhancer, 
  x_var = "prediction",   
  y_var = "ism_0_max",
  protein_name = "TEAD4"
)
plot
ggsave(
    filename = "ISM_prediction_TEAD4_2.png",
    plot = plot,
    width = 8,
    height = 6,
    dpi = 300,
    bg = "white"
  )
```

# check extreme proteins for Gradxinp
```{r}
# NFKB2
plot <- scatter_plot_with_cor2(
  data = merged_enhancer, 
  x_var = "prediction",   
  y_var = "gradxinp_0_max",
  protein_name = "NFKB2"
)
plot
ggsave(
    filename = "Gradxinp_prediction_NFKB2_1.png",
    plot = plot,
    width = 8,
    height = 6,
    dpi = 300,
    bg = "white"
  )
plot <- scatter_plot_with_cor3(
  data = merged_enhancer, 
  x_var = "prediction",   
  y_var = "gradxinp_0_max",
  protein_name = "NFKB2"
)
plot
ggsave(
    filename = "Gradxinp_prediction_NFKB2_2.png",
    plot = plot,
    width = 8,
    height = 6,
    dpi = 300,
    bg = "white"
  )

# FOXA1
plot <- scatter_plot_with_cor2(
  data = merged_enhancer, 
  x_var = "prediction",   
  y_var = "gradxinp_0_max",
  protein_name = "FOXA1"
)
plot
ggsave(
    filename = "Gradxinp_prediction_FOXA1_1.png",
    plot = plot,
    width = 8,
    height = 6,
    dpi = 300,
    bg = "white"
  )
plot <- scatter_plot_with_cor3(
  data = merged_enhancer, 
  x_var = "prediction",   
  y_var = "gradxinp_0_max",
  protein_name = "FOXA1"
)
plot
ggsave(
    filename = "Gradxinp_prediction_FOXA1_2.png",
    plot = plot,
    width = 8,
    height = 6,
    dpi = 300,
    bg = "white"
  )

# MNT
plot <- scatter_plot_with_cor2(
  data = merged_enhancer, 
  x_var = "prediction",   
  y_var = "gradxinp_0_max",
  protein_name = "MNT"
)
plot
ggsave(
    filename = "Gradxinp_prediction_MNT_1.png",
    plot = plot,
    width = 8,
    height = 6,
    dpi = 300,
    bg = "white"
  )
plot <- scatter_plot_with_cor3(
  data = merged_enhancer, 
  x_var = "prediction",   
  y_var = "gradxinp_0_max",
  protein_name = "MNT"
)
plot
ggsave(
    filename = "Gradxinp_prediction_MNT_2.png",
    plot = plot,
    width = 8,
    height = 6,
    dpi = 300,
    bg = "white"
  )
```
# ISA extreme proteins
```{r}
# VEZF1
plot <- scatter_plot_with_cor2(
  data = merged_enhancer, 
  x_var = "prediction",   
  y_var = "isa_track0",
  protein_name = "VEZF1"
)
plot
ggsave(
    filename = "ISA_prediction_VEZF1_1.png",
    plot = plot,
    width = 8,
    height = 6,
    dpi = 300,
    bg = "white"
  )
plot <- scatter_plot_with_cor3(
  data = merged_enhancer, 
  x_var = "prediction",   
  y_var = "isa_track0",
  protein_name = "VEZF1"
)
plot
ggsave(
    filename = "ISA_prediction_VEZF1_2.png",
    plot = plot,
    width = 8,
    height = 6,
    dpi = 300,
    bg = "white"
  )

# TEAD4
plot <- scatter_plot_with_cor2(
  data = merged_enhancer, 
  x_var = "prediction",   
  y_var = "isa_track0",
  protein_name = "TEAD4"
)
plot
ggsave(
    filename = "ISA_prediction_TEAD4_1.png",
    plot = plot,
    width = 8,
    height = 6,
    dpi = 300,
    bg = "white"
  )
plot <- scatter_plot_with_cor3(
  data = merged_enhancer, 
  x_var = "prediction",   
  y_var = "isa_track0",
  protein_name = "TEAD4"
)
plot
ggsave(
    filename = "ISA_prediction_TEAD4_2.png",
    plot = plot,
    width = 8,
    height = 6,
    dpi = 300,
    bg = "white"
  )

# TEF
plot <- scatter_plot_with_cor2(
  data = merged_enhancer, 
  x_var = "prediction",   
  y_var = "isa_track0",
  protein_name = "TEF"
)
plot
ggsave(
    filename = "ISA_prediction_TEF_1.png",
    plot = plot,
    width = 8,
    height = 6,
    dpi = 300,
    bg = "white"
  )
plot <- scatter_plot_with_cor3(
  data = merged_enhancer, 
  x_var = "prediction",   
  y_var = "isa_track0",
  protein_name = "TEF"
)
plot
ggsave(
    filename = "ISA_prediction_TEF_2.png",
    plot = plot,
    width = 8,
    height = 6,
    dpi = 300,
    bg = "white"
  )
```



## calculate difference
```{r}
merged_enhancer$diff_gradxinp_ism <- merged_enhancer$gradxinp_0_max - merged_enhancer$ism_0_max
merged_enhancer$diff_isa_ism <- merged_enhancer$isa_track0 - merged_enhancer$ism_0_max
```

# 10 prediction groups
```{r}
merged_enhancer <- merged_enhancer %>%
  mutate(
    prediction_class = ntile(prediction, n = 10) - 1
  ) %>%
  mutate(
    prediction_class = factor(prediction_class, levels = 0:9, ordered = TRUE)
  )

# get the range of prediction group
thresholds <- merged_enhancer %>%
  group_by(prediction_class) %>%
  summarise(
    count = n(),
    min_prediction = min(prediction),
    max_prediction = max(prediction),
    .groups = 'drop'
  )

cat("Group Distribution and Thresholds:\n")
print(thresholds, n = Inf)

```
# difference between gradxinp and ism
```{r}
library(viridis)
output_full <- "boxplot_full.png"
output_zoomed <- "boxplot_zoomed.png"

custom_theme <- theme_minimal(base_size = 12) +
  theme(
    plot.background = element_rect(fill = "white", color = NA),
    panel.background = element_rect(fill = "white", color = NA),
    panel.grid.major.x = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, size = 0.8),
    plot.title = element_text(hjust = 0.5, size = 14),
    axis.title = element_text(size = 16),   
    axis.text = element_text(size = 14)
  )

# full plot
plot_full <- ggplot(merged_enhancer, aes(x = prediction_class, 
                                         y = diff_gradxinp_ism,
                                         fill = prediction_class)) +
  geom_boxplot(alpha = 0.8, show.legend = FALSE) +
  scale_fill_viridis_d(option = "plasma") +
  labs(
    title = "Distribution of Difference between Gradxinp and ISM by Prediction Class",
    x = "Prediction Class",
    y = "Difference between Gradxinp and ISM"
  ) +
  custom_theme

# zoom
plot_zoomed <- ggplot(merged_enhancer, aes(x = prediction_class, 
                                            y = diff_gradxinp_ism,
                                            fill = prediction_class)) +
  geom_boxplot(alpha = 0.8, show.legend = FALSE, outlier.shape = NA) +  
  scale_fill_viridis_d(option = "plasma") +
  labs(
    title = "Zoomed Distribution of Difference between Gradxinp and ISM by Prediction Class",
    x = "Prediction Class",
    y = "Difference between Gradxinp and ISM"
  ) +
  coord_cartesian(ylim = c(-0.35, 0.35)) +
  custom_theme

# save
ggsave(output_full, plot_full, width = 8, height = 6, dpi = 300, bg = "white")
ggsave(output_zoomed, plot_zoomed, width = 8, height = 6, dpi = 300, bg = "white")

```
# difference between isa and ism
```{r}
output_full <- "distribution_isa.png"
output_zoomed <- "zoomed_isa.png"

# full plot
plot_full <- ggplot(merged_enhancer, aes(x = prediction_class, 
                                         y = diff_isa_ism,
                                         fill = prediction_class)) +
  geom_boxplot(alpha = 0.8, show.legend = FALSE) +
  scale_fill_viridis_d(option = "plasma") +
  labs(
    title = "Distribution of Difference between ISA and ISM by Prediction Class",
    x = "Prediction Class",
    y = "Difference between ISA and ISM"
  ) +
  custom_theme

# zoom
plot_zoomed <- ggplot(merged_enhancer, aes(x = prediction_class, 
                                            y = diff_isa_ism,
                                            fill = prediction_class)) +
  geom_boxplot(alpha = 0.8, show.legend = FALSE, outlier.shape = NA) +  # 隐藏离群值
  scale_fill_viridis_d(option = "plasma") +
  labs(
    title = "Zoomed Distribution of Difference between ISA and ISM by Prediction Class",
    x = "Prediction Class",
    y = "Difference between ISA and ISM"
  ) +
  coord_cartesian(ylim = c(-0.4, 0.4)) +
  custom_theme

ggsave(output_full, plot_full, width = 8, height = 6, dpi = 300, bg = "white")
ggsave(output_zoomed, plot_zoomed, width = 8, height = 6, dpi = 300, bg = "white")
```
# corrlation between predictcion and difference
```{r}
cor_pre_diff_grad = cor(merged_enhancer$prediction, merged_enhancer$diff_gradxinp_ism, method = "pearson")
cor_pre_diff_grad
cor_pre_diff_isa = cor(merged_enhancer$prediction, merged_enhancer$diff_isa_ism, method = "pearson")
cor_pre_diff_isa

```
## see variance between two models
```{r}
# normalization
calc_scaled_row_variance <- function(merged_enhancer) {
  data_only <- merged_enhancer[, -1]  
  abs_max <- apply(abs(data_only), 1, max)  # use max of abs
  abs_max[abs_max == 0] <- 1
  scaled_data <- data_only / abs_max  # scale
  row_vars <- apply(scaled_data, 1, var)  
  return(row_vars)
}

filtered_gradxinp_enformer <- gradxinp_enformer %>%
  filter(region %in% merged_enhancer$region)

# calculte 
deepcompre_variance <- calc_scaled_row_variance(gradxinp_deepcompre)
enformer_variance <- calc_scaled_row_variance(filtered_gradxinp_enformer)

# plot
plot_data <- data.frame(
  variance = c(deepcompre_variance, enformer_variance),
  group = rep(c("DeepCompre", "Enformer"),
              times = c(length(deepcompre_variance), length(enformer_variance)))
)


plot <- ggplot(plot_data, aes(x = group, y = variance, fill = group)) +
  geom_boxplot() +
  scale_fill_manual(values = c("steelblue", "tomato")) +
  theme_minimal() +
  labs(
    title = "Variance Comparison of Gradxinp values between two models",
    x = "Model",
    y = "Variance",
    fill = "Model"  
  ) +
    theme(
    plot.title = element_text(hjust = 0.5, size = 18),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14),
    legend.title = element_text(size = 15),
    legend.text = element_text(size = 13),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),  # 内框
    plot.background = element_rect(fill = "white", color = NA),  # 无最外层黑框
    panel.grid.major = element_line(color = "gray90")
  )

plot
ggsave(
    filename = "gradxinp_diff.png",
    plot = plot,
    width = 12,
    height = 6,
    dpi = 300,
    bg = "white"
  )

# u test
deepcompre_vals <- plot_data$variance[plot_data$group == "DeepCompre"]
enformer_vals <- plot_data$variance[plot_data$group == "Enformer"]
test_result <- wilcox.test(deepcompre_vals, enformer_vals)

print(test_result)

```











## merge two dataset
```{r}
class <- classify %>%
  filter(c_sum >= 1)
# 处理重复键 (保留第一个出现值)
merged_enhancer <- merged_enhancer %>%
  left_join(
    class %>%
      select(protein2, cooperativity_index) %>%
      distinct(protein2, .keep_all = TRUE),  # 去重
    by = c("protein" = "protein2")
  )

# 处理未匹配值 (填充默认值)
merged_enhancer <- merged_enhancer %>%
  mutate(
    cooperativity_index = coalesce(cooperativity_index, 0)  # NA替换为0
  )
```

## codenpendent and redundant motif
# ism and gradxinp
```{r}
# 数据预处理 ---------------------------------------------------------------
merged_enhancer <- merged_enhancer %>%
  mutate(
    group = case_when(
      cooperativity_index < 0.3 ~ "redundant",
      cooperativity_index > 0.7 ~ "codependent",
      TRUE ~ "unknown"  # 处理NA和中间值
    ),
    group = factor(group, levels = c("unknown", "codependent", "redundant"))
  ) %>%
  # 3. 过滤无效值
  filter(
    !is.na(diff_gradxinp_ism),
    !is.na(prediction_class)
  )

# 验证分组结果 -----------------------------------------------------------
cat("Group Distribution:\n")
merged_enhancer %>%
  count(prediction_class, group) %>%
  print(n = Inf)

# 可视化设置 -------------------------------------------------------------
group_colors <- c(
  "unknown" = "#999999",
  "codependent" = "#1b9e77",
  "redundant" = "#d95f02"
)

# 绘制箱线图 -------------------------------------------------------------
ggplot(merged_enhancer, 
       aes(x = factor(prediction_class), 
           y = diff_gradxinp_ism,
           fill = group)) +
  geom_boxplot(
    position = position_dodge(width = 0.8),  # 控制分组间距
    width = 0.7,                             # 箱体宽度
    outlier.shape = NA                       # 隐藏离群点
  ) +
  # 分面样式
  scale_fill_manual(values = group_colors) +
  labs(
    title = "Distribution of diff_gradxinp_ism by Prediction Class and Motif Type",
    x = "Prediction Class",
    y = "diff_gradxinp_ism",
    fill = "Motif Type"
  ) +
  coord_cartesian(ylim = c(-1, 1))+  # 聚焦主要数据范围
  theme_minimal() +
  theme(
    panel.grid.major.x = element_blank(),
    plot.title = element_text(hjust = 0.5, size = 14),
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    legend.position = "top"
  ) +
  # 添加统计显著性标记（可选）
  ggpubr::stat_compare_means(
    aes(group = group),
    method = "kruskal.test",
    label = "p.signif",
    show.legend = FALSE
  )

# 进阶优化版本（解决小样本显示问题）---------------------------------------
ggplot(merged_enhancer, 
       aes(x = factor(prediction_class), 
           y = diff_gradxinp_ism,
           fill = group)) +
  geom_boxplot(
    position = position_dodge(width = 0.8),
    width = 0.7,
    outlier.shape = NA
  ) +
  geom_point(
    position = position_jitterdodge(jitter.width = 0.2),
    size = 0.8,
    alpha = 0.3,
    show.legend = FALSE
  ) +
  scale_fill_manual(values = group_colors) +
  facet_wrap(~group, nrow = 1) +  # 添加分面视图
  theme(
    strip.background = element_rect(fill = "gray90"),
    strip.text = element_text(face = "bold")
  )

# according to prediction class
plot <- ggplot(merged_enhancer, 
       aes(x = factor(group),  # X 轴改为 group，使其在每个小图内分组
           y = diff_gradxinp_ism,
           fill = group)) +
  geom_boxplot(
    position = position_dodge(width = 0.8),
    width = 0.7,
    outlier.shape = NA
  ) +
  geom_point(
    position = position_jitterdodge(jitter.width = 0.2),
    size = 0.8,
    alpha = 0.3,
    show.legend = FALSE
  ) +
  scale_fill_manual(values = group_colors) +
  facet_wrap(~prediction_class, nrow = 2,scales = "free_y")+  # 根据 prediction_class 分面，每行 2 个
  theme(
    strip.background = element_rect(fill = "gray90"),
    strip.text = element_text(face = "bold")
  )
plot
ggsave(
    filename = "try.png",
    plot = plot,
    width = 12,
    height = 6,
    dpi = 300,
    bg = "white"
  )

# filter outliers
# 计算每个 prediction_class 去除离群值后的 y 轴范围
merged_enhancer_cleaned <- merged_enhancer %>%
  group_by(prediction_class) %>%
  mutate(
    Q1 = quantile(diff_gradxinp_ism, 0.25),
    Q3 = quantile(diff_gradxinp_ism, 0.75),
    IQR = Q3 - Q1,
    lower_bound = Q1 - 1.5 * IQR,
    upper_bound = Q3 + 1.5 * IQR
  ) %>%
  filter(diff_gradxinp_ism >= lower_bound & diff_gradxinp_ism <= upper_bound)

# 绘图
plot <- ggplot(merged_enhancer_cleaned, 
               aes(x = factor(group),  
                   y = diff_gradxinp_ism,
                   fill = group)) +
  geom_boxplot(
    position = position_dodge(width = 0.8),
    width = 0.7,
    outlier.shape = NA  # 不显示异常值
  ) +
  scale_fill_manual(values = group_colors) +
  facet_wrap(~prediction_class, nrow = 2, scales = "free_y") +  # 仍然允许每个面板的 y 轴独立
  theme(
    strip.background = element_rect(fill = "gray90"),
    strip.text = element_text(face = "bold")
  ) +
  theme(
    strip.text = element_text(face = "bold"),
    axis.text.x = element_text(angle = 0, hjust = 0.5)
  )

ggsave(
    filename = "prediction_class_grad_ism.png",
    plot = plot,
    width = 12,
    height = 6,
    dpi = 300,
    bg = "white"
  )


```
# calculate correlation
```{r}
# 根据 group 计算 prediction 与 diff_gradxinp_ism 和 diff_isa_ism 的相关系数
cor_by_group <- merged_enhancer %>%
  filter(!is.na(diff_gradxinp_ism), !is.na(diff_isa_ism), !is.na(prediction)) %>%
  group_by(group) %>%
  summarise(
    cor_pre_diff_grad = cor(prediction, diff_gradxinp_ism, method = "pearson"),
    cor_pre_diff_isa  = cor(prediction, diff_isa_ism, method = "pearson")
  )

print(cor_by_group)

# 对 prediction_class != "0" 的子集再分组计算
cor_by_group_non0 <- merged_enhancer %>%
  filter(prediction_class != "0",
         !is.na(diff_gradxinp_ism), !is.na(diff_isa_ism), !is.na(prediction)) %>%
  group_by(group) %>%
  summarise(
    cor_pre_diff_grad = cor(prediction, diff_gradxinp_ism, method = "pearson"),
    cor_pre_diff_isa  = cor(prediction, diff_isa_ism, method = "pearson")
  )

print(cor_by_group_non0)

```

# ism and isa
```{r}
# 绘制箱线图 -------------------------------------------------------------
ggplot(merged_enhancer, 
       aes(x = factor(prediction_class), 
           y = diff_isa_ism,
           fill = group)) +
  geom_boxplot(
    position = position_dodge(width = 0.8),  # 控制分组间距
    width = 0.7,                             # 箱体宽度
    outlier.shape = NA                       # 隐藏离群点
  ) +
  # 分面样式
  scale_fill_manual(values = group_colors) +
  labs(
    title = "Distribution of diff_isa_ism by Prediction Class and Motif Type",
    x = "Prediction Class",
    y = "diff_isa_ism",
    fill = "Motif Type"
  ) +
  coord_cartesian(ylim = c(-1, 1.5))+  # 聚焦主要数据范围
  theme_minimal() +
  theme(
    panel.grid.major.x = element_blank(),
    plot.title = element_text(hjust = 0.5, size = 14),
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    legend.position = "top"
  ) +
  # 添加统计显著性标记（可选）
  ggpubr::stat_compare_means(
    aes(group = group),
    method = "kruskal.test",
    label = "p.signif",
    show.legend = FALSE
  )

# 进阶优化版本（解决小样本显示问题）---------------------------------------
ggplot(merged_enhancer, 
       aes(x = factor(prediction_class), 
           y = diff_isa_ism,
           fill = group)) +
  geom_boxplot(
    position = position_dodge(width = 0.8),
    width = 0.7,
    outlier.shape = NA
  ) +
  geom_point(
    position = position_jitterdodge(jitter.width = 0.2),
    size = 0.8,
    alpha = 0.3,
    show.legend = FALSE
  ) +
  scale_fill_manual(values = group_colors) +
  facet_wrap(~group, nrow = 1) +  # 添加分面视图
  theme(
    strip.background = element_rect(fill = "gray90"),
    strip.text = element_text(face = "bold")
  )

# according to prediction class
plot <- ggplot(merged_enhancer, 
       aes(x = factor(group),  # X 轴改为 group，使其在每个小图内分组
           y = diff_isa_ism,
           fill = group)) +
  geom_boxplot(
    position = position_dodge(width = 0.8),
    width = 0.7,
    outlier.shape = NA
  ) +
  geom_point(
    position = position_jitterdodge(jitter.width = 0.2),
    size = 0.8,
    alpha = 0.3,
    show.legend = FALSE
  ) +
  scale_fill_manual(values = group_colors) +
  facet_wrap(~prediction_class, nrow = 2,scales = "free_y")+  # 根据 prediction_class 分面，每行 2 个
  theme(
    strip.background = element_rect(fill = "gray90"),
    strip.text = element_text(face = "bold")
  )
plot
ggsave(
    filename = "try_isa_ism.png",
    plot = plot,
    width = 12,
    height = 6,
    dpi = 300,
    bg = "white"
  )


# filter outliers
# 计算每个 prediction_class 去除离群值后的 y 轴范围
merged_enhancer_cleaned2 <- merged_enhancer %>%
  group_by(prediction_class) %>%
  mutate(
    Q1_2 = quantile(diff_isa_ism, 0.25),
    Q3_2 = quantile(diff_isa_ism, 0.75),
    IQR_2 = Q3 - Q1,
    lower_bound_2 = Q1_2 - 1.5 * IQR_2,
    upper_bound_2 = Q3_2 + 1.5 * IQR_2
  ) %>%
  filter(diff_isa_ism >= lower_bound_2 & diff_isa_ism <= upper_bound_2)

# 绘图
plot <- ggplot(merged_enhancer_cleaned2, 
               aes(x = factor(group),  
                   y = diff_isa_ism,
                   fill = group)) +
  geom_boxplot(
    position = position_dodge(width = 0.8),
    width = 0.7,
    outlier.shape = NA  # 不显示异常值
  ) +
  scale_fill_manual(values = group_colors) +
  facet_wrap(~prediction_class, nrow = 2, scales = "free_y") +  # 仍然允许每个面板的 y 轴独立
  theme(
    strip.background = element_rect(fill = "gray90"),
    strip.text = element_text(face = "bold")
  ) +
  theme(
    strip.text = element_text(face = "bold"),
    axis.text.x = element_text(angle = 0, hjust = 0.5)
  )

ggsave(
    filename = "prediction_class_isa_ism.png",
    plot = plot,
    width = 12,
    height = 6,
    dpi = 300,
    bg = "white"
  )

```





