---
title: "promoter"
output: html_document
date: "2025-02-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## read
```{r}
promoter_k562 <- read.csv("C:/Users/ruini/Desktop/thesis/motif_info_thresh_500_promoters_k562.csv", header = TRUE, sep = ",")
prediction_promoter <-read.csv("C:/Users/ruini/Desktop/thesis/predictions_promoters_k562.csv",header = TRUE)
file_path <- "C:/Users/ruini/Desktop/thesis/gene_tss.txt"
tss_data <- read.table(file_path, header = TRUE, sep = "\t", stringsAsFactors = FALSE)
colnames(tss_data)[colnames(tss_data) == "Chromosome"] <- "chromosome"
```

## set save path
```{r}
if (requireNamespace("rstudioapi", quietly = TRUE)) {
  current_path <- rstudioapi::getSourceEditorContext()$path
  if (!is.null(current_path)) {
    setwd(dirname(current_path))
    message("Working directory set to script folder: ", dirname(current_path))
  }
}
```


## data preparation
```{r}
# use abs(max) in the gradxinp for motif
promoter_k562$gradxinp_0_max <- sapply(
  strsplit(promoter_k562$gradxinp_0, ":"),
  function(x) max(abs(as.numeric(x)))
)

promoter_k562$gradxinp_0_max_orginal <- sapply(
  strsplit(promoter_k562$gradxinp_0, ":"),
  function(x) max(as.numeric(x))
)

# calculate gradxinp_min
promoter_k562$gradxinp_0_min <- sapply(
  strsplit(promoter_k562$gradxinp_0, ":"),
  function(x) min(abs(as.numeric(x)))
)

promoter_k562$gradxinp_0_min_orginal <- sapply(
  strsplit(promoter_k562$gradxinp_0, ":"),
  function(x) min(as.numeric(x))
)

# use abs(max) in the isa in base level for motif
promoter_k562$isa_0_max <- sapply(
  strsplit(promoter_k562$isa_0, ":"),
  function(x) max(as.numeric(x))
)

# use abs(max) in the ism for motif
promoter_k562$ism_0_max <- sapply(
  strsplit(promoter_k562$ism_0, ":"),
  function(x) max(as.numeric(x))
)
```

## add prediction
```{r}
merged_promoter <- merge(promoter_k562, prediction_promoter, 
                     by.x = "region", 
                     by.y = "X", 
                     all = FALSE)  
colnames(merged_promoter)[colnames(merged_promoter) == "X0"] <- "prediction"

# filter prediction score < 0
merged_promoter <- merged_promoter %>%
  filter(prediction >= 0)

merged_promoter
```

## calculate motif center
```{r}
merged_promoter <- merged_promoter %>%
  mutate(motif_center = (start + end) / 2)
```


## add TSS
```{r}
library(dplyr)
library(purrr)
library(tidyr)

find_nearest_tss <- function(motif_chr, motif_strand, motif_center) {
  motif_chr <- as.character(motif_chr)[1]
  motif_strand <- as.character(motif_strand)[1]
  motif_center <- as.numeric(motif_center)[1]

  tss_subset <- tss_data %>% filter(chromosome == motif_chr & strand == motif_strand)

  if (nrow(tss_subset) == 0) {
    return(data.frame(closest_TSS = NA_real_, tss_strand = NA_character_))
  }

  nearest_idx <- which.min(abs(tss_subset$TSS - motif_center))
  nearest_TSS <- tss_subset[nearest_idx, ]

  return(data.frame(closest_TSS = nearest_TSS$TSS, tss_strand = nearest_TSS$strand))
}

merged_promoter <- merged_promoter %>%
  mutate(
    nearest_TSS_info = pmap(
      list(chromosome, strand, motif_center),
      find_nearest_tss
    )
  ) %>%
  unnest_wider(nearest_TSS_info) %>%  
  mutate(
    relative_position = ifelse(tss_strand == "-", -(motif_center - closest_TSS), motif_center - closest_TSS),
    up_downstream = case_when(
      relative_position > 0 ~ "downstream",
      relative_position < 0 ~ "upstream",
      TRUE ~ NA_character_
    )
  )

```

## calculate threshold
```{r}
library(tidyverse)

isa_vals <- merged_promoter$isa_track0
isa_pos <- isa_vals[isa_vals > 0]
isa_neg <- isa_vals[isa_vals < 0]

# calculate 5% threshold
isa_pos_thresh <- quantile(isa_pos, 0.025, na.rm = TRUE)
isa_neg_thresh <- quantile(isa_neg, 0.975, na.rm = TRUE)  

cat("ISA threshold (left of 0):", isa_neg_thresh, "\n")
cat("ISA threshold (right of 0):", isa_pos_thresh, "\n")

# plot
ggplot(data.frame(isa = isa_vals), aes(x = isa)) +
  geom_density(fill = "skyblue", alpha = 0.5) +
  geom_vline(xintercept = c(isa_pos_thresh, isa_neg_thresh), color = "steelblue", linetype = "dashed") +
  labs(title = "ISA Track0 Distribution with ±2.5% thresholds",
       x = "isa_track0", y = "Density") +
  theme_minimal()


## for ISM
ism_vals <- merged_promoter$ism_0 %>%
  str_split(":", simplify = FALSE) %>%
  unlist() %>%
  as.numeric()

ism_pos <- ism_vals[ism_vals > 0]
ism_neg <- ism_vals[ism_vals < 0]

# calculate threshold
ism_pos_thresh <- quantile(ism_pos, 0.025, na.rm = TRUE)
ism_neg_thresh <- quantile(ism_neg, 0.975, na.rm = TRUE)
cat("ISM threshold (left of 0):", ism_neg_thresh, "\n")
cat("ISM threshold (right of 0):", ism_pos_thresh, "\n")

# plot
ggplot(data.frame(ism = ism_vals), aes(x = ism)) +
  geom_density(fill = "lightcoral", alpha = 0.5) +
  geom_vline(xintercept = c(ism_pos_thresh, ism_neg_thresh), color = "tomato", linetype = "dashed") +
  labs(title = "ISM_0 Distribution with ±2.5% thresholds",
       x = "ism_0 values", y = "Density") +
  theme_minimal()


```

## ratios for activators and repressors 
```{r}
# add ism_sign and isa_sign
merged_promoter <- merged_promoter %>%
  mutate(
    ism_sign = sapply(ism_0, function(x) {
      values <- as.numeric(unlist(strsplit(as.character(x), ":")))
      if (mean(values < -0.001, na.rm = TRUE) >= 0.7) {
        "repressor"
      } else if (mean(values > 0.001, na.rm = TRUE) >= 0.7) {
        "activator"
      } else {
        "unknown"
      }
    }),
    isa_sign = case_when(
      isa_track0 < -0.0005491227 ~ "repressor",
      isa_track0 >  0.001233367  ~ "activator",
      TRUE ~ "unknown"
    )
  )

# group by protein
protein_summary <- merged_promoter %>%
  group_by(protein) %>%
  summarise(
    ism_activator_ratio = mean(ism_sign == "activator", na.rm = TRUE),
    ism_repressor_ratio = mean(ism_sign == "repressor", na.rm = TRUE),
    
    isa_activator_ratio = mean(isa_sign == "activator", na.rm = TRUE),
    isa_repressor_ratio = mean(isa_sign == "repressor", na.rm = TRUE),
    
    ism_defined = case_when(
      ism_activator_ratio > 0.5 ~ "activator",
      ism_repressor_ratio > 0.5 ~ "repressor",
      TRUE ~ "unknown"
    ),
    isa_defined = case_when(
      isa_activator_ratio > 0.5 ~ "activator",
      isa_repressor_ratio > 0.5 ~ "repressor",
      TRUE ~ "unknown"
    ),
    .groups = "drop"
  )

print(protein_summary)
write.csv(protein_summary, "protein_summary.csv", row.names = FALSE)

```

# find position dependent repressor
```{r}
library(ggplot2)
library(ggpubr)

proteins <- unique(merged_promoter$protein)
results_list <- list()
plot_list <- list()

# for each protein
for (prot in proteins) {
  df <- merged_promoter %>%
    filter(protein == prot) %>%
    mutate(
      regulator_type = case_when(
        isa_track0 > 0.001233367 ~ "activator",
        isa_track0 < -0.0005491227 ~ "repressor",
        TRUE ~ NA_character_
      )
    ) %>%
    filter(!is.na(regulator_type))

  if (length(unique(df$regulator_type)) < 2) next

  # calculate median
  medians <- df %>%
    group_by(regulator_type) %>%
    summarize(median_position = median(relative_position), .groups = "drop")

  # u test
  test <- wilcox.test(relative_position ~ regulator_type, data = df)

  # save result
  results_list[[prot]] <- data.frame(
    protein = prot,
    activator_median = medians$median_position[medians$regulator_type == "activator"],
    repressor_median = medians$median_position[medians$regulator_type == "repressor"],
    p_value = test$p.value
  )

  # plot significant results
  if (test$p.value < 0.05) {
    p <- ggplot(df, aes(x = regulator_type, y = relative_position, fill = regulator_type)) +
      geom_violin(trim = FALSE, alpha = 0.6) +
      geom_boxplot(width = 0.1, outlier.size = 1, alpha = 0.8) +
      labs(title = paste0("Protein: ", prot, ", p = ", signif(test$p.value, 3)),
           x = "Regulator Type", y = "Relative Position") +
      theme_minimal()
    print(p)  
  }
}

results_table <- bind_rows(results_list)
print(results_table)

```

# siginificant results
```{r}
significant_proteins <- results_table %>%
  filter(p_value < 0.05)
significant_proteins
```

# filter dataset
```{r}
# filter relative_position betweeb -300 and 100 
filtered_promoter <- merged_promoter %>%
  filter(relative_position >= -300, relative_position <= 100)

proteins <- unique(filtered_promoter$protein)
results_list <- list()

# 遍历每个 protein
for (prot in proteins) {
  df <- filtered_promoter %>%
    filter(protein == prot) %>%
    mutate(
      regulator_type = case_when(
        isa_track0 > 0.001233367 ~ "activator",
        isa_track0 < -0.0005491227 ~ "repressor",
        TRUE ~ NA_character_
      )
    ) %>%
    filter(!is.na(regulator_type))

  if (length(unique(df$regulator_type)) < 2) next

  # calculate median
  medians <- df %>%
    group_by(regulator_type) %>%
    summarize(median_position = median(relative_position), .groups = "drop")

  test <- wilcox.test(relative_position ~ regulator_type, data = df)

  results_list[[prot]] <- data.frame(
    protein = prot,
    activator_median = medians$median_position[medians$regulator_type == "activator"],
    repressor_median = medians$median_position[medians$regulator_type == "repressor"],
    p_value = test$p.value
  )

  if (test$p.value < 0.05) {
    p <- ggplot(df, aes(x = regulator_type, y = relative_position, fill = regulator_type)) +
      geom_violin(trim = FALSE, alpha = 0.6) +
      geom_boxplot(width = 0.1, outlier.size = 1, alpha = 0.8) +
      labs(title = paste0("Protein: ", prot, ", p = ", signif(test$p.value, 3)),
           x = "Regulator Type", y = "Relative Position") +
      theme_minimal()
    print(p)  
  }
}

results_table <- bind_rows(results_list)
print(results_table)

significant_proteins <- results_table %>%
  filter(p_value < 0.05)
significant_proteins
write.csv(significant_proteins, "significant_proteins.csv", row.names = TRUE)
```


## distribution of activators and repressors
```{r}
# add activator / repressor label
labeled_promoter <- filtered_promoter %>%
  mutate(
    regulator_type = case_when(
      isa_track0 > 0.001233367 ~ "activator",
      isa_track0 < -0.0005491227 ~ "repressor",
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(regulator_type))


ratio_summary <- labeled_promoter %>%
  group_by(regulator_type, up_downstream) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(regulator_type) %>%
  mutate(
    total = sum(count),
    proportion = count / total
  )

print(ratio_summary)

# calculate median
median_positions <- labeled_promoter %>%
  group_by(regulator_type) %>%
  summarise(median_relative_position = median(relative_position), .groups = "drop")

print(median_positions)

# plot
plot <- ggplot(labeled_promoter, aes(x = regulator_type, y = relative_position, fill = regulator_type)) +
  geom_violin(trim = FALSE, alpha = 0.6) +
  geom_boxplot(width = 0.1, outlier.size = 1, alpha = 0.8) +
  labs(title = "Distribution of Relative Position by Regulator Type",
       x = "Regulator Type", y = "Relative Position") +
  theme_minimal()+
  scale_fill_manual(values = c("activator" = "steelblue", "repressor" = "tomato")) +
  theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),
    plot.title = element_text(hjust = 0.5, size = 18),  
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14),
    legend.position = "none", 
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_line(color = "grey90")
  )
plot
ggsave(
    filename = "distribution of activator and repressors.png",
    plot = plot,
    width = 8,
    height = 6,
    dpi = 300,
    bg = "white"
  )
```

## ratio of upstream and downstream motifs
```{r}
merged_promoter %>%
  count(up_downstream) %>%
  mutate(proportion = n / sum(n))
merged_promoter %>%
  group_by(protein, up_downstream) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(protein) %>%
  mutate(
    proportion = n / sum(n),
    total_n = sum(n)
  ) %>%
  select(-n) %>%
  pivot_wider(names_from = up_downstream, values_from = proportion, values_fill = 0) %>%
  arrange(desc(total_n))


filtered_promoter %>%
  count(up_downstream) %>%
  mutate(proportion = n / sum(n))
filtered_promoter %>%
  group_by(protein, up_downstream) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(protein) %>%
  mutate(
    proportion = n / sum(n),
    total_n = sum(n)
  ) %>%
  select(-n) %>%
  pivot_wider(names_from = up_downstream, values_from = proportion, values_fill = 0) %>%
  arrange(desc(total_n))

```

## YY1
```{r}
# filter YY1
YY1_data <- merged_promoter %>% filter(protein == "YY1")

# group by ISA score
YY1_positive <- YY1_data %>% filter(isa_track0 > 0.001233367)  
YY1_negative <- YY1_data %>% filter(isa_track0 < -0.0005491227) 
YY1_positive <- YY1_positive %>%
  mutate(group = "positive")
YY1_negative <- YY1_negative %>%
  mutate(group = "negative")

# merge dataset
merged_data <- bind_rows(YY1_positive, YY1_negative)

# plot
plot <- ggplot(merged_data, aes(x = group, y = relative_position, color = group)) +
  geom_boxplot(outlier.size = 2) +  
  labs(
    title = "Relative Position of YY1 Motifs (Positive vs Negative)",
    x = "Group",
    y = "Relative Position to TSS"
  ) +
  theme_minimal() +
  scale_color_manual(values = c("positive" = "steelblue", "negative" = "tomato")) +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 0.8),  
    plot.background = element_rect(fill = "white", color = NA),
    plot.title = element_text(hjust = 0.5, size = 18),  
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14),
    legend.position = "none"
  )
plot

ggsave(
    filename = "YY1_1.png",
    plot = plot,
    width = 8,
    height = 6,
    dpi = 300,
    bg = "white"
  )

```
# filter
```{r}
# filter -100bp to 100bp
filtered_data <- merged_data %>%
  filter(relative_position >= -100 & relative_position <= 100)

plot <- ggplot(filtered_data, aes(x=group, y=relative_position, color=group)) +
  geom_boxplot(outlier.size = 2) +
  labs(title="Relative Position of filtered YY1 Motifs (Positive vs Negative)",
       x="Group",
       y="Relative Position to TSS") +
  theme_minimal() +  
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 0.8),  
    plot.background = element_rect(fill = "white", color = NA),
    plot.title = element_text(hjust = 0.5, size = 18),  
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14),
    legend.position = "none"
  )+
  scale_color_manual(values=c("positive"="steelblue", "negative"="tomato"))
plot
ggsave(
    filename = "YY1_2.png",
    plot = plot,
    width = 8,
    height = 6,
    dpi = 300,
    bg = "white"
  )

```
## up/downsream
```{r}
# original
YY1_data <- merged_data %>%
  filter(!is.na(isa_track0), !is.na(up_downstream))  

plot <- ggplot(YY1_data, aes(x=up_downstream, y=isa_track0, color=up_downstream)) +
  geom_boxplot(outlier.size = 2) +
  labs(title="ISA scores of YY1 Motifs (Upstream vs Downstream)",
       x="Group",
       y="ISA Scores") +
  theme_minimal() +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 0.8),  
    plot.background = element_rect(fill = "white", color = NA),
    plot.title = element_text(hjust = 0.5, size = 18),  
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14),
    legend.position = "none"
  )+
  scale_color_manual(
    values = c("upstream" = "steelblue", "downstream" = "tomato"),
    name = "group"  
  )
ggsave(
    filename = "YY1_3.png",
    plot = plot,
    width = 8,
    height = 6,
    dpi = 300,
    bg = "white"
  )

# filtered
YY1_data2 <- filtered_data %>%
  filter(!is.na(isa_track0), !is.na(up_downstream))  
plot <- ggplot(YY1_data2, aes(x=up_downstream, y=isa_track0, color=up_downstream)) +
  geom_boxplot(outlier.size = 2) +
  labs(title="ISA scores of filtered YY1 Motifs (Upstream vs Downstream)",
       x="Group",
       y="ISA Scores") +
  theme_minimal() +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 0.8),  
    plot.background = element_rect(fill = "white", color = NA),
    plot.title = element_text(hjust = 0.5, size = 18),  
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14),
    legend.position = "none"
  )+
  scale_color_manual(
    values = c("upstream" = "steelblue", "downstream" = "tomato"),
    name = "group"  
  )
ggsave(
    filename = "YY1_4.png",
    plot = plot,
    width = 8,
    height = 6,
    dpi = 300,
    bg = "white"
  )
```

## TBP
```{r}
# filter TBP
TBP_data <- merged_promoter %>% filter(protein == "TBP")

# group by ISA score
TBP_positive <- TBP_data %>% filter(isa_track0 > 0.001233367)  
TBP_negative <- TBP_data %>% filter(isa_track0 < -0.0005491227) 
TBP_positive <- TBP_positive %>%
  mutate(group = "positive")
TBP_negative <- TBP_negative %>%
  mutate(group = "negative")

# merge dataset
merged_data <- bind_rows(TBP_positive, TBP_negative)

# plot
plot <- ggplot(merged_data, aes(x = group, y = relative_position, color = group)) +
  geom_boxplot(outlier.size = 2) +  
  labs(
    title = "Relative Position of TBP Motifs (Positive vs Negative)",
    x = "Group",
    y = "Relative Position to TSS"
  ) +
  theme_minimal() +
  scale_color_manual(values = c("positive" = "steelblue", "negative" = "tomato")) +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 0.8),  
    plot.background = element_rect(fill = "white", color = NA),
    plot.title = element_text(hjust = 0.5, size = 18),  
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14),
    legend.position = "none"
  )
plot

ggsave(
    filename = "TBP_1.png",
    plot = plot,
    width = 8,
    height = 6,
    dpi = 300,
    bg = "white"
  )

```
# filter
```{r}
# filter -100bp to 100bp
filtered_data <- merged_data %>%
  filter(relative_position >= -100 & relative_position <= 100)

plot <- ggplot(filtered_data, aes(x=group, y=relative_position, color=group)) +
  geom_boxplot(outlier.size = 2) +
  labs(title="Relative Position of filtered TBP Motifs (Positive vs Negative)",
       x="Group",
       y="Relative Position to TSS") +
  theme_minimal() +  
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 0.8),  
    plot.background = element_rect(fill = "white", color = NA),
    plot.title = element_text(hjust = 0.5, size = 18),  
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14),
    legend.position = "none"
  )+
  scale_color_manual(values=c("positive"="steelblue", "negative"="tomato"))
plot
ggsave(
    filename = "TBP_2.png",
    plot = plot,
    width = 8,
    height = 6,
    dpi = 300,
    bg = "white"
  )


median_positions <- filtered_data %>%
  group_by(group) %>%
  summarise(median_relative_position = median(relative_position, na.rm = TRUE))

print(median_positions)

```
## up/downsream
```{r}
# original
TBP_data <- merged_data %>%
  filter(!is.na(isa_track0), !is.na(up_downstream))  

plot <- ggplot(TBP_data, aes(x=up_downstream, y=isa_track0, color=up_downstream)) +
  geom_boxplot(outlier.size = 2) +
  labs(title="ISA scores of TBP Motifs (Upstream vs Downstream)",
       x="Group",
       y="ISA Scores") +
  theme_minimal() +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 0.8),  
    plot.background = element_rect(fill = "white", color = NA),
    plot.title = element_text(hjust = 0.5, size = 18),  
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14),
    legend.position = "none"
  )+
  scale_color_manual(
    values = c("upstream" = "steelblue", "downstream" = "tomato"),
    name = "group"  
  )
ggsave(
    filename = "TBP_3.png",
    plot = plot,
    width = 8,
    height = 6,
    dpi = 300,
    bg = "white"
  )

# filtered
TBP_data2 <- filtered_data %>%
  filter(!is.na(isa_track0), !is.na(up_downstream))  
plot <- ggplot(TBP_data2, aes(x=up_downstream, y=isa_track0, color=up_downstream)) +
  geom_boxplot(outlier.size = 2) +
  labs(title="ISA scores of filtered TBP Motifs (Upstream vs Downstream)",
       x="Group",
       y="ISA Scores") +
  theme_minimal() +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 0.8),  
    plot.background = element_rect(fill = "white", color = NA),
    plot.title = element_text(hjust = 0.5, size = 18),  
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14),
    legend.position = "none"
  )+
  scale_color_manual(
    values = c("upstream" = "steelblue", "downstream" = "tomato"),
    name = "group"  
  )
ggsave(
    filename = "TBP_4.png",
    plot = plot,
    width = 8,
    height = 6,
    dpi = 300,
    bg = "white"
  )
```


## analyse repressors
```{r}
analyze_protein_position <- function(protein_name, data, 
                                     pos_cutoff = 0.001233367, 
                                     neg_cutoff = -0.0005491227, 
                                     output_prefix = "output") {
  library(dplyr)
  library(ggplot2)

  # filter protein
  protein_data <- data %>% filter(protein == protein_name)
  
  # divide activator / repressor
  positive <- protein_data %>% filter(isa_track0 > pos_cutoff) %>% mutate(group = "positive")
  negative <- protein_data %>% filter(isa_track0 < neg_cutoff) %>% mutate(group = "negative")
  merged_data <- bind_rows(positive, negative)
  
  # plot1: full range boxplot
  plot1 <- ggplot(merged_data, aes(x = group, y = relative_position, color = group)) +
    geom_boxplot(outlier.size = 2) +
    labs(title = protein_name,
         x = "Group", y = "Relative Position to TSS") +
    theme_minimal() +
    scale_color_manual(values = c("positive" = "steelblue", "negative" = "tomato")) +
    theme(panel.border = element_rect(color = "black", fill = NA, size = 0.8),
          plot.title = element_text(hjust = 0.5, size = 18),  
          axis.title = element_text(size = 16),
          axis.text = element_text(size = 14),
          legend.position = "none", 
          plot.background = element_rect(fill = "white", color = NA))
  ggsave(paste0(output_prefix, "_1.png"), plot = plot1, width = 8, height = 6, dpi = 300, bg = "white")
  
  # filter relative position between -250bp and 50bp
  filtered_data <- merged_data %>% filter(relative_position >= -250, relative_position <= 50)
  
  # plot2: filtered range boxplot
  plot2 <- ggplot(filtered_data, aes(x = group, y = relative_position, color = group)) +
    geom_boxplot(outlier.size = 2) +
    labs(title = protein_name,
         x = "Group", y = "Relative Position to TSS") +
    theme_minimal() +
    scale_color_manual(values = c("positive" = "steelblue", "negative" = "tomato")) +
    theme(panel.border = element_rect(color = "black", fill = NA, size = 0.8),
          plot.title = element_text(hjust = 0.5, size = 18),  
          axis.title = element_text(size = 16),
          axis.text = element_text(size = 14),
          legend.position = "none",
          plot.background = element_rect(fill = "white", color = NA))
  ggsave(paste0(output_prefix, "_2.png"), plot = plot2, width = 12, height = 6, dpi = 300, bg = "white")
  
  # plot3: upstream/downstream ISA (full)
  updown_data <- merged_data %>%
    filter(!is.na(isa_track0), !is.na(up_downstream))
  
  plot3 <- ggplot(updown_data, aes(x = up_downstream, y = isa_track0, color = up_downstream)) +
    geom_boxplot(outlier.size = 2) +
    labs(title = protein_name,
         x = "Group", y = "ISA Scores") +
    theme_minimal() +
    scale_color_manual(values = c("upstream" = "tomato", "downstream" = "steelblue"), name = "group") +
    theme(panel.border = element_rect(color = "black", fill = NA, size = 0.8),
          plot.title = element_text(hjust = 0.5, size = 18),  
          axis.title = element_text(size = 16),
          axis.text = element_text(size = 14),
          legend.position = "none",
          plot.background = element_rect(fill = "white", color = NA))
  ggsave(paste0(output_prefix, "_3.png"), plot = plot3, width = 8, height = 6, dpi = 300, bg = "white")
  
  # plot4: upstream/downstream ISA (filtered)
  updown_filtered <- filtered_data %>%
    filter(!is.na(isa_track0), !is.na(up_downstream))
  
  plot4 <- ggplot(updown_filtered, aes(x = up_downstream, y = isa_track0, color = up_downstream)) +
    geom_boxplot(outlier.size = 2) +
    labs(title = protein_name,
         x = "Group", y = "ISA Scores") +
    theme_minimal() +
    scale_color_manual(values = c("upstream" = "tomato", "downstream" = "steelblue"), name = "group") +
    theme(panel.border = element_rect(color = "black", fill = NA, size = 0.8),
          plot.title = element_text(hjust = 0.5, size = 18),  
          axis.title = element_text(size = 16),
          axis.text = element_text(size = 14),
          legend.position = "none",
          plot.background = element_rect(fill = "white", color = NA))
  ggsave(paste0(output_prefix, "_4.png"), plot = plot4, width = 12, height = 6, dpi = 300, bg = "white")

  # plot5: density of relative_position (filtered)
  plot5 <- ggplot(filtered_data, aes(x = relative_position)) +
    geom_density(aes(color = group), linewidth = 1) +
    geom_density(aes(fill = group), alpha = 0.2, linewidth = 0, show.legend = FALSE) +
    scale_color_manual(values = c("positive" = "#4682B4", "negative" = "#FF6347"), name = "Group") +
    scale_fill_manual(values = c("positive" = "#4682B4", "negative" = "#FF6347"), guide = "none") +
    labs(
      title = protein_name,
      x = "Relative Position to TSS",
      y = "Density",
      color = "Group"
    ) +
    theme_minimal() +
    theme(
      panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
      panel.grid = element_blank(),
      plot.title = element_text(hjust = 0.5, size = 20),  
      axis.title = element_text(size = 18),  
      axis.text = element_text(size = 16),   
      legend.text = element_text(size = 16), 
      legend.title = element_text(size = 18),
      legend.position = c(0.05, 0.95),
      legend.justification = c(0, 1),
      legend.background = element_rect(fill = "white", color = NA),
      legend.key = element_blank()
    )

  ggsave(paste0(output_prefix, "_5_density.png"), plot = plot5, width = 10, height = 4, dpi = 300, bg = "white")

  message("Plots saved for protein: ", protein_name)
}

```
# SP1
```{r}
analyze_protein_position("SP1", merged_promoter, output_prefix = "SP1")
```
# NRF1
```{r}
analyze_protein_position("NRF1", merged_promoter, output_prefix = "NRF1")
```
# NFYA
```{r}
analyze_protein_position("NFYA", merged_promoter, output_prefix = "NFYA")
```
# IKZF1
```{r}
analyze_protein_position("IKZF1", merged_promoter, output_prefix = "IKZF1")
```
# BHLHE40
```{r}
analyze_protein_position("BHLHE40", merged_promoter, output_prefix = "BHLHE40")
```

